Selection of Polynomial Features by Genetic Algorithms

### 2013-07-29 texts/submission_01-prl/paper_spfga-review.tex
Acertos para a submissão.
mais correcções para a submissão.

\documentclass[final,authoryear,3p,times,twocolumn]{elsarticle}

@@

%
%	AMS STUFF
%
\usepackage{amsfonts}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
%

@@

%
%
%%% BEGIN DOCUMENT
%
%
\begin{document}

\begin{frontmatter}

@@

\author[ue,labmag]{Francisco Coelho\corref{fc}}
\ead{fc@di.uevora.pt}
\cortext[fc]{Corresponding author}

\author[fcul,labmag]{João Pedro Neto}
\ead{jpn@di.fc.ul.pt}

\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora}
\address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
\address[labmag]{Laboratory of Agent Modelling (LabMAg)}

\date{[{\sc draft} \today]}

\begin{abstract}
Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials. The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of optimal solution.

This paper describes a two-step method that uses genetic algorithms and linear regression to find empirical polynomial regressions. Experiments on common datasets show that, discounting the training computational effort, this method is quite competitive.
\end{abstract}

\begin{keyword}
polynomial regression \sep genetic algorithm 
\end{keyword}
\end{frontmatter}

\section{Introduction}

@@

Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models. Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}. This not only illustrates the wide spectrum of applications but, additionally, work in each one of these polynomial models uses, at some point, a genetic algorithm.

@@

A usual representation of polynomials is
$$
p\left( x_1, \ldots, x_m\right) = \sum_i \theta_i q_i
$$
where each $q_i$ is a monomial, $q_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$. For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$. The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$. For the example above
$$
A = \left[ 
\begin{array}{ccc}
1 & 0 & 0 \\
0 & 1 & 1 \\
2 & 0 & 1
\end{array}
\right] \sim 
\left[ 
\begin{array}{ccc}
x_1 \\
x_2 x_3 \\
x_1^2 x_3
\end{array}
\right]
$$
where each row defines a monomial and each column represents a variable. Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
%$$
%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
%\begin{array}{ccc}
%x_1 \\
%x_2 x_3 \\
%x_1^2 x_3
%\end{array}
%\right].
%$$

@@

The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $Q$ a set of polynomials. For each polynomial $p\in Q$ let $\mathcal{P}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
%
\begin{eqnarray*}
\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{P}_p,\mathcal{D}\right)
\end{eqnarray*}
%
by solving the first problem. With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $Q$ until a reasonable approximation of a local minimum is found. 
%
\begin{algorithm}[tb]
\begin{algorithmic}
\Function{GAPoly}{$D,pop_0,\epsilon$}
	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
	\While{$err > \epsilon$}
		\State $pop \gets \Call{IterateGA}{pop}$
		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$
		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
	\EndWhile
	\State\Return $\Call{First}{pop}$
\EndFunction
\end{algorithmic}
\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials. At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
\end{algorithm} 
%
Notice that the properties of \acp{GA} and linear regression entail that
the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
%

@@

Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
\begin{verbatim}
110 - 011,000,001 ; 000,000,111 ;
      001,001,000 ; 110,010,101
\end{verbatim}

@@

\begin{figure}[tb]\begin{center}
\includegraphics[width=0.48\textwidth]{figure_2.pdf}

@@

\end{center}\end{figure}

>>>>

\documentclass[review,preprint,authoryear,12pt]{elsarticle}

%
%	LINE NUMBERING FOR REVISION
%
\usepackage{lineno}
%

@@

%
%	AMS STUFF
%
\usepackage{amsfonts}
%\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
%

@@

%
%
%%% BEGIN DOCUMENT
%
%
\begin{document}
\begin{frontmatter}

@@

\author[ue,labmag]{Francisco ~Coelho\corref{fc}}
\ead{fc@di.uevora.pt}
\cortext[fc]{Corresponding author}

\author[fcul,labmag]{João ~Pedro ~Neto}
\ead{jpn@di.fc.ul.pt}

\address[ue]{Dept. Informática, Universidade de Évora, Rua Romão Ramalho 58, 7000-671 Évora}
\address[fcul]{Dept. Informática, Faculdade de Ciências da Universidade de Lisboa, Campo Grande 1749-016 Lisboa}
\address[labmag]{Laboratory of Agent Modelling (LabMAg)}
%
%\date{[{\sc draft} \today]}
%
\begin{abstract}
Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials. The use of genetic algorithms to find polynomial models is decades old but still poses challenges due to the complexity of the search and different definitions of optimal solution.

This paper describes a two-step method that uses genetic algorithms and linear regression to find empirical polynomial regressions. Experiments on common datasets show that, discounting the training computational effort, this method is quite competitive.
\end{abstract}
%
\begin{keyword}
polynomial regression \sep genetic algorithm \sep feature extraction \sep dimensionality reduction 
\end{keyword}
\end{frontmatter}
%
%	Line numbers for revision
%
%
\linenumbers
%
\section{Introduction}

@@

Polynomials, one of the most studied subjects in mathematics, generalize li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models. Applications include colorimetric calibration \citep{mendes2005adaptive}, explicit formul\ae\ for turbulent pipe flows \citep{davidson1999method}, computational linguistics \citep{sanchez2009obtaining} and more recently, analytical techniques for cultural heritage materials \citep{csefalvayova2010use}, liquid epoxy molding process \citep{chan2011modeling}, B-spline surface reconstruction \citep{galvez2012iterative}, product design \citep{chan2012development} or forecasting cyanotoxins presence in water reservoirs \citep{garcia2013hybrid}. These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.

@@

An usual representation of polynomials is
$$
p\left( x_1, \ldots, x_m\right) = \sum_i \theta_i q_i
$$
where each $q_i$ is a monomial, $q_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$. For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $q_1 = x_1, q_2 = x_2 x_3$ and $q_3 = x_1^2 x_3$, coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$. The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$.

For the example above the matrix of monomials is
$$
A = \left[ 
\begin{array}{ccc}
1 & 0 & 0 \\
0 & 1 & 1 \\
2 & 0 & 1
\end{array}
\right] \sim 
\left[ 
\begin{array}{ccc}
x_1 \\
x_2 x_3 \\
x_1^2 x_3
\end{array}
\right]
$$
where each row defines a monomial and each column represents a variable. Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
%$$
%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
%\begin{array}{ccc}
%x_1 \\
%x_2 x_3 \\
%x_1^2 x_3
%\end{array}
%\right].
%$$

@@

The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $Q$ a set of polynomials. For each polynomial $p\in Q$ let $\mathcal{P}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
%
\begin{eqnarray*}
\phi_p &=& \min_\Theta J_{\textrm{fit}}\left(\Theta;\mathcal{P}_p,\mathcal{D}\right)
\end{eqnarray*}
%
by solving the first problem. With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $Q$ until a reasonable approximation of a local minimum is found. 
%
\begin{algorithm}[tb]
\begin{algorithmic}
\Function{GAPoly}{$D,pop_0,\epsilon$}
	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
	\While{$err > \epsilon$}
		\State $pop \gets \Call{IterateGA}{pop}$
		\State $pop \gets \Call{Sort}{pop, key = J}$\Comment{Sort population by regression error}
		\State $err \gets J\left(\Call{First}{pop}\right)$
	\EndWhile
	\State\Return $\Call{First}{pop}$
\EndFunction
\end{algorithmic}
\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials. The role of the linear regression step is to produce a fitting error that sorts the population. At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
\end{algorithm} 
%
Notice that the properties of \acp{GA} and linear regression entail that
the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the fitness function, encapsulated in the \ac{rmse} function $J_{\textrm{fit}}$.
%

@@

Let's see an example: consider polynomial $x_1^3 x_3 + x_3^7 + x_1 x_2$ with $k = 4, m = 3$ and $d=3$. One possible encoding would be:
\begin{center}
\texttt{110 - 011,000,001 ; 000,000,111 ; 001,001,000 ; 110,010,101}
\end{center}

@@

\begin{figure*}[tb]\begin{center}
\includegraphics[width=0.98\textwidth]{figure_2.pdf}

@@
